%I want to thank Dr. Truszczynski, other professors on my committee board, my colleagues,
%my parents, my wife and son, and my friends.

Time flies.  I cannot believe that my graduate school career is coming to the end.
It almost feels yesterday when I started my Ph.D. program in computer science at the
University of Kentucky (UK) in Fall 2010.
Looking back and connecting the dots in the past six years,
I owe much gratitude to the wonderful individuals riding along with me from start
to end of this journey.

First and foremost, I want to thank my advisor and mentor, Dr. Miroslaw Truszczynski,
for his continuous support of my research, for always being there whenever
I had questions, and for his patience and immense knowledge.
I am thankful to his guidance throughout the years that has helped me
grow to not only a better computer science researcher, but a better
writer and speaker in writing technical papers, including this dissertation,
and presenting research results at numerous venues.

Besides, I am very grateful to the rest of my research committee:
Dr. Samson Cheung, Dr. Jane Hayes and Dr. Jerzy Jaromczyk,
for their insightful and valuable comments that encouraged me
to expand my research from various perspectives.
These research directions provide me with a promising
starting point for my next endeavor as an assistant
professor.

Next, my sincere thanks goes to all the professors at UK
whom I am fortunately to have taken classes with.
I am also thankful to all the classmates, among
other peers, for their helpful and, more often than
not, cheerful discussions during the years.

Finally, none of this dissertation would have been possible
without the support of my beloved parents, my father
Zhaoling Liu and my mother Ping Liu, who had implanted
in me the great value of hard working and optimism, long
before I started graduate school.
I also thank my wife Xiaozhen Zhang
for her unconditional love and caring through all
the ups and downs, and my son Adam for
making me laugh and mature gradually as a caregiver
and as a father.
