\singlespacing
\hypersetup{
    colorlinks,%
    citecolor=black,%
    filecolor=black,%
    linkcolor=black,%
    urlcolor=black 
    %urlcolor=mygreylink     % can put red here to better visualize the links
}
\urlstyle{same}
\definecolor{mygrey}{gray}{.85}
\definecolor{mygreylink}{gray}{.40}
\textheight=9.0in
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Adjust margins
\addtolength{\oddsidemargin}{-0.375in}
\addtolength{\evensidemargin}{0.375in}
\addtolength{\textwidth}{0.5in}
\addtolength{\topmargin}{-.375in}
\addtolength{\textheight}{0.75in}

%-----------------------------------------------------------
%Custom commands
\newcommand{\resitem}[1]{\item #1 \vspace{-2pt}}
\newcommand{\resheading}[1]{{\large \colorbox{mygrey}{\begin{minipage}{\textwidth}{\textbf{#1 \vphantom{p\^{E}}}}\end{minipage}}}}
\newcommand{\ressubheading}[4]{
\begin{tabular*}{6.5in}{l@{\extracolsep{\fill}}r}
		\textbf{#1} & #2 \\
		\textit{#3} & \textit{#4} \\
\end{tabular*}\vspace{-6pt}}

\newcommand{\ressubsubheading}[2]{
\begin{tabular*}{6.5in}{l@{\extracolsep{\fill}}r}
		\textit{#1} & \textit{#2} \\
\end{tabular*}\vspace{-6pt}}
\renewcommand{\baselinestretch}{0.9}
%-----------------------------------------------------------

\newcommand{\mywebheader}{
\begin{tabular*}{6.5in}{l@{\extracolsep{\fill}}r}
	\textbf{\LARGE Xudong Liu} & 
		\href{http://www.cs.uky.edu/~liu/}{http://www.cs.uky.edu/$\sim$liu/} $\bullet$ 
		liu@cs.uky.edu\\
	\end{tabular*}
\\
\vspace{0.1in}}

% CHANGE HEADER SOURCE HERE
\mywebheader

%%%%%%%%%%%%%%%%%%%%%%
\resheading{Research Interests}
%I am looking for a 2015 summer intern position in software development engineering.
My research focuses on exciting topics in artificial intelligence and social science, including
preferences (i.e., preference modeling, learning and reasoning), 
data-driven and computational social science, 
knowledge representation and reasoning, 
decision theory, data mining, and machine learning.
\vspace{3pt}

%%%%%%%%%%%%%%%%%%%%%%
\resheading{Education}
			\ressubheading{University of Kentucky}{USA}
			{Doctor of Philosophy, computer science
				}
			{Aug. 2010 -- May 2016}\\  \vspace{6pt}
			{GPA:\textbf{4.00}/4.00 \\ Advisor: Dr. Miroslaw Truszczynski}  \vspace{-6pt}
				%{ \small
				\begin{itemize}
					%\resitem{\textbf{Available for internship: July 23, 2009; Graduation follows internship}}
					\resitem{\textbf{Courses}: 
							Modern Operating Systems, 
							Numerical Analysis, 
							Algorithm Design,
							Distributed Operating System, 
							Satisfiability and Equivalence Checking, 
							Database Systems, 
							Computer Networks, 
							Networks Security, 
							Preferences in AI and Decision Theory, 
							Principles of Constraint Satisfaction,
							Comparative Decision Making, 
							Preparing Future Faculty, and Grant Writing.}
				\end{itemize}
				%}
%\begin{comment}
			\ressubheading{Harbin Institute of Technology}{China}
				{Bachelor of Engineering, software engineering}{Aug. 2006 -- July 2010}\\ \vspace{6pt}
				{GPA:3.56/4.00 \\ Advisor: Prof. Yushan Sun}
%\end{comment}
\vspace{3pt}


%%%%%%%%%%%%%%%%%%%%%%
\resheading{Employment}
			\ressubheading{{R\&D Intern}}{Palo Alto Research Center (PARC), USA}
			{Supervisor: Dr. Christian Fritz}{Jun. 2015 -- Aug. 2015}
				{ 
				\begin{itemize}
					\resitem{Conducted research on and developed system modules for representing 
						and reasoning about user constraints and preferences in 
						trip planning.}
				\end{itemize}
				}
			\ressubheading{{Graduate Research Assistant}}{University of Kentucky, USA}
			{Advisor: Dr. Miroslaw Truszczynski}
			{Aug. 2010 -- May 2015}
				{
				\begin{itemize}
					\resitem{Conducted research on logic-based knowledge representation formalisms, 
						studied and built
						tools for representing and reasoning about constraints and preferences in artificial intelligence.}
				\end{itemize}
				}
			\ressubheading{{Graduate Teaching Assistant}}{University of Kentucky, USA}
			{Advisors: Dr. Truszczynski, Dr. Pike and Dr. Moore}
			{Aug. 2010 -- May 2015}
				{
				\begin{itemize}  \itemsep -2pt
					\resitem{CS215 - Introduction to Program Design and Problem Solving: leading lab
						sessions, grading assignments, preparing exam questions and solutions,
						and holding office hours.}
					\resitem{CS375 - Logic and Theory of Computing: grading assignments, 
						preparing assignment solutions,
						and holding office hours.}
					\resitem{CS463G - Introduction to Artificial Intelligence: guest instructor,
						on knowledge representation and reasoning, e.g., propositional logic, and
						first-order logic.}
				\end{itemize}
				}
			\ressubheading{{Undergraduate Teaching Assistant}}{Harbin Institute of Technology, China}
			{Advisor: Prof. Yushan Sun}
			{Aug. 2008 -- May 2010}
				{
				\begin{itemize}  \itemsep -2pt
							\resitem{Compilers and J2EE, including
								leading lab sessions and grading assignments.}
				\end{itemize}
				}
			\ressubheading{{Undergraduate Intern}}{Information Security Lab, Harbin Institute of Technology, China}
			{Advisor: Prof. Yushan Sun}
			{Aug. 2009 -- May 2010}
				{
				\begin{itemize}  \itemsep -2pt
							\resitem{Used PHP, MySQL and Apache Tomcat to implement modules of a management system.}
				\end{itemize}
				}
\vspace{3pt}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\resheading{Professional Services}
	\begin{description} \itemsep -3pt
		\item[Student member:] {Association for the Advancement of Artificial Intelligence (AAAI)}
		\item[Student volunteer:] {29th AAAI Conference on Artificial Intelligence (AAAI-15)}
		\item[Program committee:] {IJCAI-16, IJCAI-13}
		\item[Paper reviewer:] {JAIR, AAAI-14, ISAIM-14}
		\item[Local arrangement committee:] {ADT-15, LPNMR-15, ICLP-11, NonMon@30-10}
	\end{description} % End Skills list
\vspace{3pt}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\resheading{Refereed Publications}
\ressubheading{Conferences:}{}
{}{} \vspace{-0.4cm}
\begin{enumerate}  \itemsep -3pt
	\item \textbf{Xudong Liu}.
		\textit{Modeling, Learning and Reasoning with Qualitative Preferences}.
		In Proceedings of the 4th International Conference on Algorithmic Decision Theory (ADT),
		volume 9346, pages 587-592, 2015. Springer
	\item \textbf{Xudong Liu} and Miroslaw Truszczynski.
		\textit{Reasoning with Preference Trees over Combinatorial Domains}.
		In Proceedings of the 4th International Conference on Algorithmic Decision Theory (ADT),
		volume 9346, pages 19-34, 2015. Springer
	\item \textbf{Xudong Liu} and Miroslaw Truszczynski.
		\textit{Learning Partial Lexicographic Preference Trees over Combinatorial Domains}.
		In Proceedings of the 29th AAAI Conference on Artificial Intelligence (AAAI),
		pages 1539-1545, 2015. AAAI Press
	\item \textbf{Xudong Liu} and Miroslaw Truszczynski.
		\textit{Aggregating Conditionally Lexicographic Preferences Using Answer Set Programming Solvers}.
		In Proceedings of the 3rd International Conference on Algorithmic Decision Theory (ADT),
		volume 8176, pages 244-258, 2013. Springer
	\item Matthew Spradling, Judy Goldsmith, {\bf Xudong Liu}, Chandrima Dadi and Zhiyu Li.
		\textit{Roles and Teams Hedonic Game}.
		In Proceedings of the 3rd International Conference on Algorithmic Decision Theory (ADT), 
		volume 8176, pages 351-362, 2013. Springer
\end{enumerate}

\ressubheading{Workshops:}{}
{}{} \vspace{-0.4cm}
\begin{enumerate}  \itemsep -3pt
	\item \textbf{Xudong Liu} and Miroslaw Truszczynski.
		\textit{Preference Trees: A Language for Representing and Reasoning about Qualitative Preferences}.
		In Proceedings of the 8th AAAI Multidisciplinary Workshop on Advances in Preference 
		Handling (MPREF), pages 55-60, 2014. AAAI Press
	%\item {\bf Xudong Liu} and Miroslaw Truszczynski.
	%			\textit{Aggregating Conditionally Lexicographic Preferences
	%			Using Answer Set Programming Solvers}.
	%			In the 7th IJCAI Multidisciplinary Workshop on Advances in Preference Handling (MPREF), 2013.
	%\item Matthew Spradling, Judy Goldsmith, {\bf Xudong Liu}, Chandrima Dadi and Zhiyu Li.
	%			\textit{Roles and Teams Hedonic Game}.
	%			In the 7th IJCAI Multidisciplinary Workshop on Advances in Preference Handling (MPREF), 2013.
\end{enumerate}

\ressubheading{Abstract:}{}
{}{} \vspace{-0.4cm}
\begin{enumerate}  \itemsep -3pt
	\item {\bf Xudong Liu}. \textit{Aggregating Lexicographic Preference Trees Using Answer Set Programming: Extended Abstract}. 
In 23rd International Joint Conference on Artificial Intelligence Doctoral Consortium (IJCAI DC), 2013.
\end{enumerate}

\vspace{3pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\resheading{Technical Talks}

\begin{enumerate}  \itemsep -3pt
	\item \textit{Personalization in Trip Planning}. Oral.
		At the Keeping Current Seminar, Department of Computer Science,
		University of Kentucky, 2015.
	\item \textit{Preference and Social Choice over Combinatorial Domains}. Oral. 
		At the 1st ADT/LPNMR Doctoral Consortium (ADT/LPNMR-DC-15),
		Lexington, Kentucky, USA, 2015.
	\item \textit{Modeling, Learning and Reasoning with Qualitative Preferences}. Poster. 
		At the 1st ADT/LPNMR Doctoral Consortium (ADT/LPNMR-DC-15),
		Lexington, Kentucky, USA, 2015.
	\item \textit{Reasoning with Preference Trees over Combinatorial Domains}. Oral.
		At the 4th International Conference on Algorithmic Decision Theory (ADT-15),
		Lexington, Kentucky, USA, 2015.
	\item \textit{On Personalizability and Extensibility of Multi-Modal Trip Planning}. Oral.
		At Dialog, Palo Alto Research Center, Palo Alto, California, USA, 2015.
	\item \textit{Constraints and Preferences in Multi-modal Trip Planning}. Poster.
		At the PARC Summer Intern Poster Session, Palo Alto Research Center, 
		Palo Alto, California, USA, 2015.
	\item \textit{Learning Partial Lexicographic Preference Trees over Combinatorial Domains}. 
		Oral (ratio: 12\%=238/1991).
		At the 29th AAAI Conference on Artificial Intelligence (AAAI-15), 
		Austin, Texas, USA, 2015.
	\item \textit{Preference Trees: A Language for Representing and Reasoning about Qualitative Preferences}. Oral.
		At the 8th Multidisciplinary Workshop on Advances in Preference Handling (MPREF-14), 
		Quebec City, Canada, 2014.
	\item \textit{Aggregating Conditionally Lexicographic Preferences Using Answer Set Programming Solvers}. Oral.
		At the 3rd International Conference on Algorithmic Decision Theory (ADT-13),
		Universite lebre de Bruxelles, Belgium, 2013.
	\item \textit{Reasoning About Lexicographic Preferences Over Combinatorial Domains}.
		At the Keeping Current Seminar, Department of Computer Science, 
		University of Kentucky, 2013.
	\item \textit{Roles and Teams Hedonic Game}. Oral.
		At the 7th Multidisciplinary Workshop on Advances in Preference Handling (MPREF-13),
		Tsinghua University, Beijing, China, 2013.
	\item \textit{Aggregating Conditionally Lexicographic Preferences Using Answer Set Programming Solvers}. Oral.
		At the 7th Multidisciplinary Workshop on Advances in Preference Handling (MPREF-13),
		Tsinghua University, Beijing, China, 2013.
	\item \textit{Aggregating Lexicographic Preference Trees Using Answer Set Programming: Extended Abstract}. Poster.
		At the 23rd International Joint Conference on Artificial Intelligence 
		Doctoral Consortium (IJCAI-DC-13), Tsinghua University, Beijing, China, 2013.
	\item \textit{Answer Set Programming using Gringo/Clasp}. Oral.
		At the Keeping Current Seminar, Department of Computer Science,
		University of Kentucky, 2011.
\end{enumerate}
\vspace{3pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\resheading{Honors and Awards}
{\sl Verizon Fellowship}  \hfill Fall 2015 - Spring 2016\\
{\sl Graduate Teaching Assistantship}  \hfill Fall 2014 - Spring 2015, Fall 2012 - Spring 2013\\
{\sl AAAI-15 Student Volunteer and Scholarship Award}  \hfill Jan. 2015\\
{\sl International Student Tuition Scholarship}  \hfill Jan. 2015\\
{\sl Nominee of the Dissertation Year Fellowship}  \hfill Dec. 2014\\
{\sl Harrison D. Brailsford Graduate Scholarship}  \hfill Oct. 2014\\
{\sl Kentucky Opportunity Fellowship Awards}  \hfill July 2013 - June 2014\\
{\sl Nominee of the ACM Award for Outstanding Teaching Assistant}  \hfill 2013\\
{\sl NSF Student Travel Award}  \hfill Aug. 2013\\
{\sl IJCAI-13 Travel Grant Award}  \hfill Aug. 2013\\
{\sl Graduate Research Assistantship}  \hfill Fall 2010 - Spring 2013\\
{\sl Daniel R. Reedy Quality Achievement Fellowship} \hfill Aug. 2010 - May 2013\\
{\sl UK Student Travel Funding Awards}  \hfill 2013 (IJCAI, ADT), 2014 (AAAI)\\
{\sl Chinese National Endeavor Scholarship} \hfill Fall 2008\\
{\sl Outstanding Student Scholarships} \hfill Fall 2006 -  Spring 2010
\vspace{3pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\resheading{Technical Skills}
	\begin{description} \itemsep -3pt
		\item[Programming languages:] { C/C++,
			C\#, Java, Matlab, Python, Perl, SQL, PHP, HTML}
		\item[System and software:] { 
			Linux/Macintosh/Windows, Qt, Git, Answer Set
			Programming tools, Microsoft Visual Studio, Eclipse, MySql, Microsoft SQL Server
		}
	\end{description} % End Skills list


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\resheading{Selected Projects}
%
%\begin{description}  \itemsep -3pt
%\item[Preference Tree Learning System:] 
%	{ \footnotesize Given a set of pairwise preferences, called examples, 
%	between combinatorial objects,
%	the system learns a partial lexicographic preference tree, preferably small in size,
%	that are consistent with all (or preferably many) given examples.
%	[C++, matlab]}
%
%\item[Logic and SAT Based Social Choice System:] 
%	{ \footnotesize Given votes expressed as lexicographical preference trees, the system
%	computes either a winner that is an alternative with the highest score according
%	to a voting rule (Borda, k-approval, etc), or an alternative whose score is at
%	least some given threshold value. The two problems are modeled as ASP and WPM
%	programs.
%	[C++, clingo, clingcon, toulbar2; Supported by NSF grant IIS-0913459]}
%
%\item[Voting-based Multi-player Gaming Matching System:]
%	{ \footnotesize Given individual preferences from players on team 
%	compositions and individual strategies within a composition, 
%  the system computes a good match of teams of players that 
%	optimize the players’ preferences to some degree.
%	[C++, python, matlab; Partially supported by NSF grant IIS-0913459]}
%
%\item[Preferences Modeling and Optimization:] 
%	{ \footnotesize A user is asked to provide preferences with importance levels. A preference is
%	expressed as a propositional formula or an answer set optimization preference
%	rule for reasoning in the settings of possibilistic logic, leximin ordering, discrimin
%	ordering and Pareto dominance. Optimal solutions are computed.
%	[C++, flex\&bison, gidl/minisatid, clingo; Supported by NSF grant IIS-0913459]}
%
%\item[QoS IP Routing Decision Making Circuits:] 
%	{ \footnotesize Led a programming team to design and implement a GENI network on top of
%	ProtoGENI, where a source node established a route from itself to a destination
%	node by sending \textit{link state advertisement} packets and using Dijkstra's
%	algorithm to compute the shortest path.
%	[C++, ProtoGENI]}
%
%\item[Database System:] 
%	{ \footnotesize Implemented a database system from scratch that handles transactions such as
%	queries, aggregates, insertion and deletion.
%	[Java, Eclipse]}
%\end{description}
%\vspace{3pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\resheading{References}
{\bf Dr. Miroslaw Truszczynski}, Professor\\
Department of Computer Science,
University of Kentucky \\
329 Rose Street,
Lexington, KY 40506 USA \\
Office Phone: (859) 257-6738\\
Email: mirek@cs.engr.uky.edu
\vspace{0.2cm}

{\bf Dr. Christian Fritz}, Area Manager\\
System Sciences Lab,
Palo Alto Research Center \\
3333 Coyote Hill Road,
Palo Alto, CA 94304 USA \\
Office Phone: (650) 812-4876 \\
Email: cfritz@parc.com
\vspace{0.2cm}

{\bf Dr. Victor Marek}, Professor\\
Department of Computer Science,
University of Kentucky \\
329 Rose Street,
Lexington, KY 40506 USA \\
Office Phone: (859) 257-3496\\
Email: marek@cs.uky.edu
\vspace{0.2cm}

{\bf Dr. Yi Pike}, Lecturer\\
Department of Computer Science,
University of Kentucky \\
329 Rose Street,
Lexington, KY 40506 USA \\
Office Phone: (859) 218-5769\\
Email: yipike@cs.uky.edu
